---
title: "Analizing Forest Fire Data"
Author: Matias L. Conde
dataset: forestfires.csv
output: html_notebook
---

### In this Notebook we are going to try to answer the following questions: 
-  During which months are forest fires most common? 
-  On which days of the week are forest fires most common?
-  One showing the number of forest fires occuring during each month
-  Another showing the number of forest fires occurring on each day of the week

```{r}
library(readr)
library(ggplot2)
library(dplyr)
library(purrr)
```
#### Reading the dataset:
```{r}
forest_fires <- read_csv("/home/matias/Documents/repos/Data-Science-Notebooks/Analyzing Forest Fire Data/forestfires.csv")

View(forest_fires)

forest_fires <- forest_fires %>%
  mutate(month = factor(month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")), 
         day = factor(day, levels = c("sun","mon","tue","wed","thu","fri","sat")))
```
#### DISTRIBUTION OF NUMBER OF FIRES PER DAY AND PER MONTH:
```{r}
fires_per_day <- forest_fires %>%
  group_by(day) %>%
  summarize(`Number of Forest-Fire` = n()) # %>%

fires_per_month <- forest_fires %>%
  group_by(month) %>%
  summarize(`Number of Forest-Fire` = n()) # %>%

ggplot(data = fires_per_day) + 
  aes(x = day, y = `Number of Forest-Fire`) + 
  geom_bar(stat = "identity", colour="gold", fill="gold", alpha=0.3) + 
  theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))	+ 
  labs(title = "Number of Forest Fire per Day")

ggplot(data = fires_per_month) + 
  aes(x = month, y = `Number of Forest-Fire`) + 
  geom_bar(stat = "identity", colour="gold", fill="gold", alpha=0.3) + 
  theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))	+ 
  labs(title = "Number of Forest Fire per Month")

```
#### DISTRIBUTION OF ANOTHER VARIABLES PER DAY AND PER MONTH. 


 FFMC: Fine Fuel Moisture Code index from the FWI system: 18.7 to 96.20
 DMC: Duff Moisture Code index from the FWI system: 1.1 to 291.3
 DC: Drought Code index from the FWI system: 7.9 to 860.6
 ISI: Initial Spread Index from the FWI system: 0.0 to 56.10
 temp: Temperature in Celsius degrees: 2.2 to 33.30
 RH: Relative humidity in percentage: 15.0 to 100
 wind: Wind speed in km/h: 0.40 to 9.40
 rain: Outside rain in mm/m2 : 0.0 to 6.4
 area: The burned area of the forest (in ha): 0.00 to 1090.84 

```{r}
vars <- c("FFMC","DMC","DC","ISI","temp","RH","wind","rain")


create_box_plots_day <- function(x,y) {
  ggplot(data = forest_fires) + 
    aes_string(x = x, y = y) + 
    geom_boxplot(colour="darkorange", fill="papayawhip") + 
    labs(title = "Variables per Day of the Week") + 
    theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))
}

create_box_plots_month <- function(x,y) {
  ggplot(data = forest_fires) + 
    aes_string(x = x, y = y) + 
    geom_boxplot(colour="firebrick2", fill="white") + 
    labs(title = "Variables per Month") + 
    theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))
}

```
#### BOX PLOTS PER DAY
```{r}
x_var_day = names(forest_fires)[4]
y_var_day = names(forest_fires)[5:12]

map2(x_var_day, y_var_day, create_box_plots_day)
```
#### BOX PLOTS PER MONTH
```{r}
x_var_month = names(forest_fires)[3]
y_var_month = names(forest_fires)[5:12]

map2(x_var_month, y_var_month, create_box_plots_month)
```


####  The area variable contains data on the number of hectares of forest that burned during the forest fire. 
####  Perhaps we can use this variable as an indicator of the severity of the fire. 
####  The idea is that worse fires will result in a larger burned area. 

####  Let's create scatter plots to see what can be learned about relationships between 
####  The area burned by a forest fire and the following variables: 

```{r}
create_scatter <- function(x,y) {
  ggplot(data = forest_fires) + 
    aes_string(x = x, y = y) + 
    geom_point(alpha = 0.3,  colour="deepskyblue") + 
    labs(title = "Burned area vs other variables") + 
    theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))
}

x_var2 = names(forest_fires[5:12])
y_var2 = names(forest_fires[13])

map2(x_var2, y_var2, create_scatter)
```

#### The result is that most points are clustered around the bottom of the plots. A histogram of the area variable values clearly illustrates the distribution and why the scatter plots look the way they do: 

```{r}
ggplot(data = forest_fires) + 
  aes(x = area) + 
  geom_histogram()
```
#### We can see that a lot of values are arround of the zero. Then we can subset data into three categories: 0 value of area, 10-20 y 20 - 50

```{r}
area_not_null <- forest_fires %>%
  filter(area > 1 & area <100)

create_scatter2 <- function(x,y) {
  ggplot(data = area_not_null) + 
    aes_string(x = x, y = y) + 
    geom_point(alpha = 0.3, colour="deepskyblue") + 
    labs(title = "Burned area vs other variables") + 
    theme(panel.background = element_rect(fill = "white"), plot.title = element_text(hjust = 0.5))
}

x_var2 = names(forest_fires[5:12])
y_var2 = names(forest_fires[13])

map2(x_var2, y_var2, create_scatter2)
```
